Electron and ion acceleration by relativistic shocks: particle-in-cell simulations

### 2018-01-12 papers/Physica2017/Romansky_Physica2017.tex
paper

The simulations are one-dimensional and have the following parameters of the initial flow: Lorentz factor $\gamma = 1.5$, the number densities $n_e = 10^{-4} \rm{cm}^{-3}$, $n_p = 10^{-4} \rm{cm}^{-3}$ , the temperature, $T_e = T_p = 5\cdot10^8 \rm{K}$, the magnetic field $B = 10^{-4} \rm{G}$, the full size of the box $L = 2\cdot10^{12} \rm{cm}$, the number of cells $N=2\cdot10^4$. The electron mass is increase to $m_e = \frac{m_p}{20}$ to enlarge electron scales of  the plasma, it is common approach in PIC simulations see e.g. {\cite{Sironi2011}} to reduce required computational recources. The full time of simulation is $T = 5000 {\omega_p}^{-1}$, where $\omega_p$ is electron plasma frequency.These values allow one to define the dimensionless parameter magnetization $\sigma = B^2/4\pi\gamma (n_p m_p + n_e m_e) c^2 = 0.003$. Inclination angle $\theta$ is the angle between the flow velocity and the magnetic field. We present results for the particle spectrum in several simulations at various $\theta$. Stable shocks formed in all simulations and the total energy deviation is less thanof the system conserved to better than $3\%$, which indicates that the code provides a correct model of the shock waves.

@@

One can see in Figures \ref{protons} and \ref{electrons} the particle spectra for various inclination angles $\theta$. The spectra consist of three parts - the narrow peak of cold but fast moving initial flow, the wide peak of hot downstream and the non-thermal accelerated component. Figures show that the spectrum of accelerated particles strongly depends on he inclination angle $\theta$. If $\theta$ is less than the critical value, defined by the equation $c\cdot \cos(\theta_{crit})=v_{shock}$, where all values are measured in the upstream rest frame, particles can escape from the front and cross it several times to gain more energy (see, e.g., \cite{Pelletier2017}), if $\theta$ is greater, shock becomes superluminal - particles need to move along the field lines faster than light to escape from the front, and acceleration is suppressed. It is difficult to evaluate the critical angle \textit{a priori}, because the compression relation of the shock which will be created in the simulation is unknown. However the estimate for the case of a strong wave (the compression relation equals $4$) provides $\theta_{crit}=45^{\circ}$ in the downstream frame, which is consistent with the results of our simulations. The small-scale oscilations of spectra are produced by numerical errors and should not be taken into account.

@@

\section*{References}
\begin{thebibliography}{20}
	\bibitem{Bell1978} Bell A R 1978 \textit{MNRAS} \textbf{182} 147
	\bibitem{Blandford1978} Blandford R D and Ostriker J P 1978 \textit{ApJ} \textbf{221} L29 
	\bibitem{Bykov2014} Bykov A M, Ellison D C, Osipov S M and Vladimirov A E 2014 \textit{ApJ} \textbf{789} Issue 2, 137
	\bibitem{Berezhko2003} Berezhko E G, Ksenofontov L T and V{\"o}lk H J  2003 \textit{A}{\&}\textit{A} \textbf{412} L11
	\bibitem{Uchiyama2007} Uchiyama Y, Aharonian F A, Tanaka T, Takahashi T and Maeda Y 2007 \textit{Nature} \textbf{449} 576
	\bibitem{Romansky2016} Romansky V I, Bykov A M, Osipov S M and Gladilin P E \textit{Journal of Physics: CS}
	\bibitem{Lapenta2006} Lapenta G, Brackbill J U and Ricci P 2006 \textit{Phys. Plasmas} \textbf{13} 055904
	\bibitem{Noguchi2007} Noguchi K, Tronci C, Zuccaro G and Lapenta G 2007 \textit{Phys. Plasmas} \textbf{14} 042308
	\bibitem{Ellison2013} Ellison D C, Warren D C and Bykov A M 2013 \textit{ApJ} \textbf{776} Issue 1, 46
	\bibitem{Pelletier2017} Pelletier G, Bykov A M, Ellison D C and Lemoine M 2017 \textit{Space Science Reviews} \textbf{207} Issue 1-4, pp. 319-360
	\bibitem{Sironi2011} Sironi L and Spitkovsky A 2011 \textit{ApJ} \textbf{741} Issue 1, 39
\end{thebibliography}
\end{document}

>>>>

The simulations are one-dimensional and have the following parameters of the initial flow: Lorentz factor $\gamma = 1.5$, the number densities $n_e = 10^{-4} \rm{cm}^{-3}$, $n_p = 10^{-4} \rm{cm}^{-3}$ , the temperature, $T_e = T_p = 5\cdot10^8 \rm{K}$, the magnetic field $B = 10^{-4} \rm{G}$, the full size of the box $L = 2\cdot10^{12} \rm{cm}$, the number of cells $N=2\cdot10^4$. The electron mass $m_e$ is increase by $100$ times to enlarge electron scales of  the plasma, it is common approach in PIC simulations see e.g. {\cite{Sironi2011}} to reduce required computational recources. The full time of simulation is $T = 5000 {\omega_p}^{-1}$, where $\omega_p$ is electron plasma frequency.These values allow one to define the dimensionless parameter magnetization $\sigma = B^2/4\pi\gamma (n_p m_p + n_e m_e) c^2 = 0.003$. Inclination angle $\theta$ is the angle between the flow velocity and the magnetic field. We present results for the particle spectrum in several simulations at various $\theta$. Stable shocks formed in all simulations and the total energy of the system conserved to better than $3\%$, which indicates that the code provides a correct model of the shock waves.

@@

One can see in Figures \ref{protons} and \ref{electrons} the particle spectra for various inclination angles $\theta$. The spectra consist of three parts - the narrow peak of cold but fast moving initial flow, approximated by Maxwell distribution, shifted on the initial momentum, the wide peak of hot downstream, approximated by Maxwell-J{\"u}ttner distribution, and the non-thermal accelerated component. The small-scale oscilations of spectra are produced by numerical errors and should not be taken into account. Figures show that the spectrum of accelerated particles strongly depends on he inclination angle $\theta$. If $\theta$ is less than the critical value, defined by the equation $c\cdot \cos(\theta_{crit})=v_{shock}$, where all values are measured in the upstream rest frame, particles can escape from the front and cross it several times to gain more energy (see, e.g., \cite{Pelletier2017}), if $\theta$ is greater, shock becomes superluminal - particles need to move along the field lines faster than light to escape from the front, and acceleration is suppressed. It is difficult to evaluate the critical angle \textit{a priori}, because the compression relation of the shock which will be created in the simulation is unknown. However the estimate for the case of a strong wave (the compression relation equals $4$) provides $\theta_{crit}=45^{\circ}$ in the downstream frame, which is consistent with the results of our simulations.

@@

\section*{References}
\begin{thebibliography}{20}
	\bibitem{Bell1978} Bell A R 1978 \textit{MNRAS} \textbf{182} 147
	\bibitem{Blandford1978} Blandford R D and Ostriker J P 1978 \textit{ApJ} \textbf{221} L29 
	\bibitem{Bykov2014} Bykov A M, Ellison D C, Osipov S M and Vladimirov A E 2014 \textit{ApJ} \textbf{789} Issue 2, 137
	\bibitem{Berezhko2003} Berezhko E G, Ksenofontov L T and V{\"o}lk H J  2003 \textit{A}{\&}\textit{A} \textbf{412} L11
	\bibitem{Uchiyama2007} Uchiyama Y, Aharonian F A, Tanaka T, Takahashi T and Maeda Y 2007 \textit{Nature} \textbf{449} 576
	\bibitem{Romansky2016} Romansky V I, Bykov A M, Osipov S M and Gladilin P E 2017 \textit{Journal of Physics: CS} \textbf{929} id 012014 
	\bibitem{Lapenta2006} Lapenta G, Brackbill J U and Ricci P 2006 \textit{Phys. Plasmas} \textbf{13} 055904
	\bibitem{Noguchi2007} Noguchi K, Tronci C, Zuccaro G and Lapenta G 2007 \textit{Phys. Plasmas} \textbf{14} 042308
	\bibitem{Ellison2013} Ellison D C, Warren D C and Bykov A M 2013 \textit{ApJ} \textbf{776} Issue 1, 46
	\bibitem{Pelletier2017} Pelletier G, Bykov A M, Ellison D C and Lemoine M 2017 \textit{Space Science Reviews} \textbf{207} Issue 1-4, pp. 319-360
	\bibitem{Sironi2011} Sironi L and Spitkovsky A 2011 \textit{ApJ} \textbf{741} Issue 1, 39
\end{thebibliography}
\end{document}
